%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Medium Length Professional CV
% LaTeX Template
% Version 2.0 (8/5/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Rishi Shah 
%
% Important note:
% This template requires the resume.cls file to be in the same directory as the
% .tex file. The resume.cls file provides the resume style used for structuring the
% document.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[10pt]{resume} % Use the custom resume.cls style

\usepackage[left=0.5in,top=0.5in,right=0.5in,bottom=0.5in]{geometry} % Document margins
\newcommand{\tab}[1]{\hspace{.05\textwidth}\rlap{#1}}
\newcommand{\itab}[1]{\hspace{0em}\rlap{#1}}
\name{Thomas J. Brailey} % Your name
\address{UK: +44 7786 345 443 \\ US: +1 (858) 610-2437} % Your address
\address{thomasjbrailey@gmail.com \\ https://github.com/tjbrailey} % Your phone number and email

\begin{document}
	
	%----------------------------------------------------------------------------------------
	%	EDUCATION SECTION
	%----------------------------------------------------------------------------------------
	
	\begin{rSection}{Education}
		
		{\bf University of Oxford} \hfill {2022 - 2024} 
		\\ MPhil, Comparative Government
		
		{\bf University of California, San Diego} \hfill {2016 - 2020} 
		\\ Bachelor of Science, Political Science/Data Analytics \hfill {\em summa cum laude}
		%\\ Coursework: Spatial Data Analysis, Causal Inference, Applied Data Analysis, Game Theory, Research Design
		
	\end{rSection}
	
	%----------------------------------------------------------------------------------------
	%	WORK EXPERIENCE SECTION
	%----------------------------------------------------------------------------------------
	
	\begin{rSection}{Research Experience}
		
		\begin{rSubsection}{Payments and Governance Research Program}{2020-2022}{Pre-Doctoral Fellow}{}
		\item Conducted exploratory data analysis and data visualization across several research projects.
		\item Coordinated the transition of all code to GitHub and introducing RAs/GSRs to version control. 
		\item Edited manuscripts and preparing presentations for government leaders and other stakeholders.
		\end{rSubsection}
		
		\begin{rSubsection}{Center for the Study of African Political Economy}{2018-2020}{Research Assistant}{}
		\item Independently developed a large-N database of election statistics across all recognized states.
		\item Used R to combine and reshape multiple datasets, visualize the data, and perform statistical analysis.
		\item Manually inputted missing data from newspapers and reports to ensure completeness of data. 
		\end{rSubsection}
		
		\begin{rSubsection}{Center for Peace and Security Studies}{2018-2020}{Research/Lab Assistant}{}
		\item Used R to conduct quantitative analyses on large-N datasets.
		\item Generated graphs, tables, and figures in R and LaTeX for research papers and conference presentations.
		\item Wrote briefs on my data analysis and code and presented findings. 
			
		\end{rSubsection}
		
		%\begin{rSubsection}{Revelle College}{2018-2019}{Resident Assistant}{}
		%\item Fostered an inclusive community within Revelle through event programming and outreach.
		%\item Acted as a dependable resource for residents regarding academics, emotional support, and social justice. 
		%\item Received regular training on issues including autism awareness, sexual assault, and mental health issues. 
		%\end{rSubsection}
		
		%\begin{rSubsection}{Scripps Institution of Oceanography}{2017-2018}{Government Relations Intern}{}
		%	\item Reviewed and researched congressional testimonies as they related to SIO projects.
		%	\item Compiled summaries and created databases of government policy and legislation.
		%	\item Offered and accepted paid position at end of internship.
		%\end{rSubsection}
		
	\end{rSection}
	
	%----------------------------------------------------------------------------------------
	%	Publications
	%----------------------------------------------------------------------------------------
	
	\begin{rSection}{Publications and Working Papers}\itemsep -5pt
		
		\begin{rSubsection}{As Research Assistant}{}{}{}
		\item Karthik Muralidharan, Paul Niehaus, and Sandip Sukhtankar. \textit{Identity Verification Standards in Welfare Programs: Experimental Evidence from India}. Forthcoming at The Review of Economics and Statistics. October 2022. 
		\item Karthik Muralidharan, Paul Niehaus, and Sandip Sukhtankar. \textit{General equilibrium effects of (improving) public employment programs: experimental evidence from India}. Forthcoming at Econometrica. October 2022. 
		\item Stefan Faridani and Paul Niehaus. \textit{Rate-optimal linear estimation of average global effects}. Working Paper. September 2022.
		\item J. Andres Gannon. \textit{One if by Land, and Two if by Sea: Cross-Domain Contests and the Escalation of International Crises}. International Studies Quarterly. September 2022.
		\item Nikhar Gaikwad and Gareth Nellis. \textit{Overcoming the Political Exclusion of Migrants: Theory and Experimental Evidence from India}. American Political Science Review. June 2021. 
		\item Erik Gartzke, Jon R. Lindsay, J. Andres Gannon, and Peter Schram. \textit{The Shadow of Deterrence: Why capable actors engage in contests short of war}. Under Review. December 2020. 
		\item Kerry Chavez, and J. Andres Gannon. \textit{Cruisin’ for a Bruisin’: The Effects of US Cruise Missile Strikes Since the Gulf War}. Work in Progress. December 2020.
		\end{rSubsection}
	
	\begin{rSubsection}{As (Co) Author}{}{}{}
		\item Thomas Brailey. \textit{Provisions of Power-Sharing: Assessing the Impacts of Segmental Autonomy}. Working Paper. April 2021.
		\item Inbok Rhee and Thomas Brailey. \textit{Targeting the core or swing?: Campaign Strategy during the 2017 Kenyan Presidential Election}. Work in Progress. June 2021.
	\end{rSubsection}
	
	\end{rSection}
	
	%----------------------------------------------------------------------------------------
	%	Conferences
	%----------------------------------------------------------------------------------------
	\begin{rSection}{Conference Invites}
		\itemsep -5pt
		\item American Political Science Association (Virtual, 2021).
		\item International Studies Association (Virtual, 2021).
		\item Midwestern Political Science Association (Chicago, 2020, cancelled due to COVID-19).
		\item International Studies Association (Honolulu, 2020, cancelled due to COVID-19).
		\item Southern California Research Conference (San Diego, 2019).
		\item Western Political Science Association (San Diego, 2019).
	\end{rSection}
	
	%----------------------------------------------------------------------------------------
	%	AWARDS
	%----------------------------------------------------------------------------------------
	\begin{rSection}{Scholarship and Awards}
		\itemsep -5pt
		\item Merton College Research Grant 2023.
		\item Clarendon Scholar 2022-2024.
		\item BITSS Catalyst Grant Awardee 2021.
		\item Sanford Lakoff Award for the Most Outstanding Senior Thesis 2020.
		\item Pi Sigma Alpha 2017-2020. 
		\item Provost Honors 2016-2020. 
		\item Ernest C. Mort Leadership Excellence Award 2018. 
		\item Outstanding Resident 2017.
	\end{rSection}
	
	%----------------------------------------------------------------------------------------
	% Volunteer Experience
	%----------------------------------------------------------------------------------------
	\begin{rSection}{Volunteer Experience} \itemsep -5pt
		\item Volunteer data scientist for DataKind UK.
		\item Co-founder of the Political Science Student Association at UCSD.
		\item Employment Assistant at the International Rescue Committee.
		\item Tutor for San Diego's Refugee Tutoring Program.
		%\item Campaign Intern and Data Analyst for Mike Levin’s congressional campaign.
		%\item Data assistant to the national coordinator of the Citizens Climate Lobby, UK. 
		%\item Member of the campus-wide Community Standards Board.
	\end{rSection}

	%----------------------------------------------------------------------------------------
	% Training
	%----------------------------------------------------------------------------------------
	\begin{rSection}{Training} \itemsep -5pt
		\item Oxford Spring School for Advanced Research Methods (Machine Learning, 2023).
	\end{rSection}

	%----------------------------------------------------------------------------------------
	% Skills
	%----------------------------------------------------------------------------------------
	\begin{rSection}{Skills} \itemsep -5pt
		\begin{rSubsection}{Spoken Languages}{}{}
			\item English (Native), French (Intermediate), Arabic (Beginner).
		\end{rSubsection}
		\begin{rSubsection}{Technical Languages}{}{}
			\item Proficient: ArcGIS, Git, \LaTeX, R, SPSS, STATA.
			\item Beginner: Python, SQLite/SpatiaLite, Tableau.
		\end{rSubsection}
		\begin{rSubsection}{Infrastructures}{}{}
			\item AWS, Microsoft Office.
		\end{rSubsection}
	\end{rSection}
	
	\begin{rSection}{References}
		References available upon request. 
	\end{rSection}
	
\end{document}
