---
title: "Finding a country"
author: "Tom Brailey"
date: "2021-10-04"
categories: ["R"]
tags: ["R", "functions", "web-scraping"]
---

# Introduction

While working on my blog post about foreign paramilitaries in Mali, I realized that the data I was using had the name of the militia group, but not the country of origin. I tasked myself with the challenge of writing a function that, given a input phrase or name of a group, returns the corresponding country of origin. I couldn't really gauge whether this would be a predominantly regex-based or web-scraping-based issue, but as I explored various avenues of solution, I realized that it would be a mixture of both. 

After a couple of days of trial and error, I created the function. Here it is below:

```{r data, eval=TRUE, include=FALSE}

# Let's load our essential packages
library(magrittr)
library(ggplot2)
library(rvest)
library(httr)
library(data.table)

# Source our country-finder function
source(paste0(here::here(), "/content/find_country.R"))

# Set up googlesheets connection
googlesheets4::gs4_deauth()

# Election data
elec  <- 
  googlesheets4::read_sheet(
    ss = "https://docs.google.com/spreadsheets/d/14RxWGtmXL-ZAhYZlnvarn5CBqIp0DcGr7wkbatY8oWs/edit#gid=0", 
    na = "NA")

# ACLED data
acled <- 
  read.csv2(
    paste0(
      here::here(),
      "/public/posts/2019-01-29-2022-02-02-Eastern_Africa-Middle_Africa-Northern_Africa-Southern_Africa-Western_Africa.csv"), 
    sep = ";") 

# Creating subsets for testing (town/village guesses)
set.seed(1)
testing_dat_1 <- acled %>% 
  dplyr::select(country, admin2) %>% 
  dplyr::filter(admin2 != "") %>% 
  dplyr::distinct() %>% 
  dplyr::slice_sample(n = 200)

# Creating subsets for testing (actor guesses)
set.seed(1)
testing_dat_2 <- acled %>%
  dplyr::mutate(
    actor1_location_temp = stringr::str_extract_all(
      string = actor1, pattern = "\\([^()]+\\)", simplify = FALSE),
    actor1_location_temp = stringr::str_remove_all(
      string = actor1_location_temp, 
      pattern = "\\(|\\)|\\-|\\d+|\\W\\W|character"),
    actor1 = stringr::str_remove_all(
      string = actor1, pattern = "\\([^()]+\\)")) %>% 
  dplyr::mutate(
    actor1_location = countrycode::countryname(
      sourcevar = actor1_location_temp)) %>% 
  dplyr::filter(!is.na(actor1_location)) %>% 
  dplyr::select(actor1_location, actor1) %>% 
  dplyr::distinct() %>% 
  dplyr::slice_sample(n = 200)

# Get a list of countries from acled
countries_acled <- readRDS(paste0(here::here(), "/public/posts/acled_countries.rds"))
countries_acled <- countries_acled$Country
countries_acled <- sub(" \\(.*| of | & | and | the |  ", " ", countries_acled)
countries_acled <- sub(" \\(.*| of | & | and | the |  ", " ", countries_acled)
countries_acled <- unique(countries_acled)
countries_acled <- sort(countries_acled, decreasing = TRUE)
countries_acled

```

```{r load, include=TRUE, eval=TRUE}

find_country <- function(name, root = TRUE) {
  
  # Display keywords you are searching for
  #print(paste0("Searching for keyphrase: ", name))
  
  # Generate URL
  url  <- URLencode(paste0("https://www.google.com/search?q=", name#, " wikipedia"
                           ))
  page <- rvest::read_html(url)
  
  # Extract all relevant links
  nodes <- rvest::html_nodes(page, "a")
  links <- rvest::html_attr(nodes,"href")
  
  # Iterate over possible links to find country
  for(x in 1:length(links)){
    
    # Extract first link of the search results
    link <- links[startsWith(links, "/url?q=")][x]
    
    # Clean link
    link <- sub("^/url\\?q\\=(.*?)\\&sa.*$","\\1", link)
    link <- sub("25","", link)
    link <- sub("%E2%2580%2593", "–", link)
    link <- sub("%25C3%25A9", "é", link)
    
    # Tell user which link you're trying 
    #print(paste0("Trying link ", link))
    
    # Enter site
    site <- try(
      expr = link %>% 
        httr::GET(., httr::timeout(10)) %>% 
        rvest::read_html(.), 
      silent = TRUE)
    
    # If list of links is exhausted, break and go to next keyphrase
    if(sum(class(site) == "try-error") == 1 & x == length(links)){
      country <- NA 
      break
    }
    
    # If link times out or is bad, go to next link
    if(sum(class(site) == "try-error") == 1){
      next
    }
    
    # Pull all words from first few paragraphs of site 
    text <- try(
      expr = site %>% 
      html_nodes("p") %>% 
      html_text() %>% 
      .[nchar(.) > 150] %>%
      strsplit(split = " ") %>% 
      unlist(.) %>%
      stringr::str_replace_all(., pattern = "of|the", "") %>%
      stringr::str_replace_all(., pattern = "\\W+|\\d+", " ") %>%
      stringr::str_c(., collapse = " ") %>%
      stringr::str_squish(.),
      silent = TRUE)
    
    # If link error, go to next link
    if(sum(class(text) == "try-error") == 1){
      next
    }
    
    # If data has 0 observations, go to next link
    if(length(text) == 0){
      next
    }
    
    # Pull key country names from text
    matching_dat <- 
      tibble::as_tibble(
        unlist(
          stringr::str_match_all(
            text, 
            stringr::str_c(
              countries_acled, collapse="|")))) %>% 
      dplyr::filter(!is.na(value)) %>%
      dplyr::distinct(value)
    
    # Save matched cases as a vector
    partial_match <- matching_dat$value
    
    # Loop through strings to find country match
    i <- 1
    for(i in 1:length(partial_match)){
      
      pm <- paste(partial_match[1:i], collapse = " ")
      
      # If we get a match, break inner loop
      if(sum(stringr::str_detect(pattern = countries_acled, string = pm)) > 0){
        country <- 
          countries_acled[stringr::str_detect(
            pattern = pm, 
            string = countries_acled)]
        break
      }
    }
    
    # If we get a match, break outer loop 
    if(exists("country")){
      break
    }
  }
  
  # Print likely match and return value 
  #print(paste0("Most likely country: ", country, "."))
  Sys.sleep(2)
  return(country)
}

```

And here is an example of it in action:

```{r test, eval=FALSE,include=TRUE}

find_country(name = "Wagner group")

```

In words, this function takes as its input a string of something you want to learn more about. In the example, I wanted to learn where the "Wagner Group", a paramilitary militia operating in Mali and the Central African Republic, originates. The function takes this string, and searches for the most relevant Wikipedia page. If there is no corresponding Wikipedia page, then the function defers to the next most relevant match. Then, the function pulls the first few paragraphs of text from the page and matches any countries mentioned with a pre-generate list of countries. If the country is the first-mentioned or most frequently mentioned, the function assumes this to be the target country and returns it as a string---Russia, as in the example above. Let's try another: Say I wanted to remind myself where Timbuktu is...

```{r test 2, eval=FALSE,include=TRUE}

find_country(name = "Timbuktu")

```
Aha! What a coincidence---we were just talking about Mali...

This function is primitive, and I want to add more complexity to it over time. There are numerous issues with the assumption that the "first" hit is always going to correspond to the target country, but let's empirically test how well this function performs in it's current state. Let's subset the ACLED data to include the country, and the town or village of that country. We can then input the village name into the function and compare how often it figures out the actual country.

```{r empirical test, eval=FALSE,include=TRUE}

testing_dat_1$guess_country <- lapply(X = testing_dat_1$admin2, FUN = find_country)

```

Okay, now we've guessed the countries, let's see how well we did:

```{r empirical test 2, eval=FALSE,include=TRUE}

plot_dat <- testing_dat_1 %>% 
  tidyr::unnest(cols = guess_country) %>% 
  dplyr::group_by(country, admin2) %>% 
  dplyr::mutate(num_guess = dplyr::n()) %>%
  dplyr::mutate(
    guess_country = dplyr::case_when(
      guess_country == "Democratic Republic Congo" ~ "Democratic Republic of Congo",
      TRUE ~ guess_country)) %>%
  dplyr::mutate(
    correct = ifelse(
      stringr::str_detect(
        pattern = guess_country, 
        string = country) == TRUE, 1, 0),
    x = ifelse(num_guess > 1 & max(correct) == 1, 1, NA)) %>% 
  dplyr::filter(!(num_guess > 1 & max(correct) == 1 & x == 1))

pct_correct <- round(sum(plot_dat$correct==1, na.rm = TRUE) / nrow(plot_dat), 2)

ggplot(
  plot_dat, 
  mapping = aes(x = factor(correct))) + 
  geom_histogram(stat = "count") + 
  labs(
    x = "Was the guess correct?", 
    y = "Frequency",
    subtitle = paste0("Correct ratio: ", pct_correct)) + 
  theme_classic()

```

Hmm, 67%, not bad. Let's try something more challenging:

```{r empirical test 3, eval=FALSE,include=TRUE}

testing_dat_2$guess_country <- lapply(X = testing_dat_2$actor1, FUN = find_country)

```

```{r empirical test 2, eval=FALSE,include=TRUE}

plot_dat2 <- testing_dat_2 %>% 
  tidyr::unnest(cols = guess_country) %>% 
  dplyr::group_by(actor1_location, actor1) %>% 
  dplyr::mutate(num_guess = dplyr::n()) %>%
  dplyr::mutate(
    guess_country = dplyr::case_when(
      guess_country == "Democratic Republic Congo" ~ "Democratic Republic of Congo",
      TRUE ~ guess_country)) %>%
  dplyr::mutate(
    correct = ifelse(
      stringr::str_detect(
        pattern = guess_country, 
        string = actor1_location) == TRUE, 1, 0),
    x = ifelse(num_guess > 1 & max(correct) == 1, 1, NA)) %>% 
  dplyr::filter(!(num_guess > 1 & max(correct) == 1 & x == 1))

pct_correct <- round(sum(plot_dat2$correct == 1, na.rm = TRUE) / nrow(plot_dat2), 2)

ggplot(
  plot_dat2, 
  mapping = aes(x = factor(correct))) + 
  geom_histogram(stat = "count") + 
  labs(
    x = "Was the guess correct?", 
    y = "Frequency",
    subtitle = paste0("Correct ratio: ", pct_correct)) + 
  theme_classic()

```
